0/1 Knapsack

### 2015-03-13 Programming Project C/01knapsack.tex
final update

%% BEGIN ABSTRACT
\begin{abstract}
The present study investigates two implementations for a 0/1 Knapsack problem, a greedy solution, and a dynamic solution. The dynamic solution will always result in the correct answer but takes $\theta{n*m}$ time to complete. The greedy solution will generally result in a reasonable answer, and even sometimes the correct answer, and take $\theta{n}$ time to complete. Which solution to use will depend on what is most important to the problem at hand, speed, or accuracy.

@@

%% BEGIN BACKGROUND
\section{Background}
Unfortunately, solving a 0/1 Knapsack problem often times proves difficult because in order to find the best solution, all possible solutions must be attempted, a method known as brute force. It is not practical to brute force a problem of this kind because it would take $ \theta{n!} $ to solve, which is unacceptable for large sets of numbers. However, there are other methods beside brute force that allow for the correct or 'good-enough' solution to be found. One of these methods is known as dynamic programming, which trades off speed for space. With dynamic programming the correct solution can be found in $ \theta{n*m} $ time, where n is the amount of items and m is the size of the knapsack. Another method for solving this type of problem is called a greedy algorithm, which works by making the best decision on which item to take locally. A greedy solution to the 0/1 Knapsack problem traditionally runs in $ \theta{n} $ time. Unfortunately a greedy solution is not always correct and therefore highly unreliable.

@@

%% BEGIN EXPERIMENTAL ANALYSIS
\section{Experimental Analysis}
\begin{figure}[!]
\begin{center}
\includegraphics[scale=.55]{test-results.pdf}
\end{center}
\captionof{figure}{Java Implementation of Greedy and Dynamic Solutions to 0/1 Knapsack Run-Time Analysis}
\label{fig:runtime}
\end{figure}
The insertion sort, greedy solution and dynamic solution demonstrated in Algorithm 1, 2 and 3 was implemented in Java and executed on an HP SpectreXT TouchSmart with a 4 core Intel i7 Ivybridge processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit with 8GB of DDR3 RAM.
The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta{n^2} $ where $ \theta{} $ represents the asymptotically tightly bound running time. \cite{textbook}

The greedy solution also follows a similar $ \theta{n^2} $ run time, see Figure 1. This is because an insertion sort was chosen to implement as the sorting algorithm used in the greedy solution. An insertion sort was chosen for the sorting algorithm because, while it is not the fastest, requires less axillary space. While run time is an important factor, the space required to run the algorithm is also very important, and should not be neglected. As seen in Table 2, the dynamic solution unavoidably required too much axillary space and Java gave an error when values of $n$ became increasingly large. It also important to note that the data graphed in Figure 1 was executed while the machine executed superfluous commands such as checking for system updates. To aid this issue, averages of 10 trails were taken and used in Figure 1.

The run time for the dynamic solution as seen in figure one $ \theta{n*c} $ where n is the number of elements and c is the size of the knapsack. In this implementation for values of $n$ greater than or equal to 100, a value of $2n$ was used for $c$. That would make the run time of the specific implementation of the dynamic solution $2n^2$ but after dropping constants becomes $ \theta{n^2} $. In Figure 1, the reason that the dynamic solution takes so much more time to complete is because there is a constant of 2 for $n^2$ which can be attributed to doubling the real time it took to return the results. Because the dynamic solution works by storing answers originally found in axillary space, the values of $n$ cannot be very high without causing a Java error, as seen in Table 2.

@@

%% BEGIN CONCLUSION
\section{Conclusion}
Although in this specific implementation, both the dynamic solution and the greedy solution had a run-time of $\theta{n^2}$, it can be possible for the dynamic solution to have a run-time of $\theta{n*c}$ and the greedy solution to have a run-time of $\theta{n}$. The best algorithm to use for a given problem is really a subjective decision and depends on two things, 1. how accurate does the answer to the problem really need to be, and 2. what is the time requirement to find out what the answer is.

>>>>

%% BEGIN ABSTRACT
\begin{abstract}
The present study investigates two implementations for a 0/1 Knapsack problem, a greedy solution, and a dynamic solution. The dynamic solution will always result in the correct answer but takes $\theta(n*m)$ time to complete. The greedy solution will generally result in a reasonable answer, and even sometimes the correct answer, and take $\theta(n)$ time to complete. Which solution to use will depend on what is most important to the problem at hand, speed, or accuracy.

@@

%% BEGIN BACKGROUND
\section{Background}
Unfortunately, solving a 0/1 Knapsack problem often times proves difficult because in order to find the best solution, all possible solutions must be attempted, a method known as brute force. It is not practical to brute force a problem of this kind because it would take $ \theta(n!) $ to solve, which is unacceptable for large sets of numbers. However, there are other methods beside brute force that allow for the correct or 'good-enough' solution to be found. One of these methods is known as dynamic programming, which trades off speed for space. With dynamic programming the correct solution can be found in $ \theta(n*m) $ time, where n is the amount of items and m is the size of the knapsack. Another method for solving this type of problem is called a greedy algorithm, which works by making the best decision on which item to take locally. A greedy solution to the 0/1 Knapsack problem traditionally runs in $ \theta(n) $ time. Unfortunately a greedy solution is not always correct and therefore highly unreliable.

@@

%% BEGIN EXPERIMENTAL ANALYSIS
\section{Experimental Analysis}
\begin{figure}[!]
\begin{center}
\includegraphics[scale=.55]{test-results.pdf}
\end{center}
\captionof{figure}{Java Implementation of Greedy and Dynamic Solutions to 0/1 Knapsack Run-Time Analysis}
\label{fig:runtime}
\end{figure}
The insertion sort, greedy solution and dynamic solution demonstrated in Algorithm 1, 2 and 3 was implemented in Java and executed on an HP SpectreXT TouchSmart with a 4 core Intel i7 Ivybridge processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit with 8GB of DDR3 RAM.
The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta(n^2) $ where $ \theta() $ represents the asymptotically tightly bound running time. \cite{textbook}

The greedy solution also follows a similar $ \theta(n^2) $ run time, see Figure 1. This is because an insertion sort was chosen to implement as the sorting algorithm used in the greedy solution. An insertion sort was chosen for the sorting algorithm because, while it is not the fastest, requires less axillary space. While run time is an important factor, the space required to run the algorithm is also very important, and should not be neglected. As seen in Table 2, the dynamic solution unavoidably required too much axillary space and Java gave an error when values of $n$ became increasingly large. It also important to note that the data graphed in Figure 1 was executed while the machine executed superfluous commands such as checking for system updates. To aid this issue, averages of 10 trails were taken and used in Figure 1.

The run time for the dynamic solution as seen in figure one $ \theta(n*c) $ where n is the number of elements and c is the size of the knapsack. In this implementation for values of $n$ greater than or equal to 100, a value of $2n$ was used for $c$. That would make the run time of the specific implementation of the dynamic solution $2n^2$ but after dropping constants becomes $ \theta(n^2) $. In Figure 1, the reason that the dynamic solution takes so much more time to complete is because there is a constant of 2 for $n^2$ which can be attributed to doubling the real time it took to return the results. Because the dynamic solution works by storing answers originally found in axillary space, the values of $n$ cannot be very high without causing a Java error, as seen in Table 2.

@@

%% BEGIN CONCLUSION
\section{Conclusion}
Although in this specific implementation, both the dynamic solution and the greedy solution had a run-time of $\theta(n^2)$, it can be possible for the dynamic solution to have a run-time of $\theta(n*c)$ and the greedy solution to have a run-time of $\theta(n)$. The best algorithm to use for a given problem is really a subjective decision and depends on two things, 1. how accurate does the answer to the problem really need to be, and 2. what is the time requirement to find out what the answer is.
